A comparison between the latest Prophet forecasting model and the ARIMA model: An evaluation of BTC/ZAR predictability

### 2017-10-27 Draft/Write-up/Draft.Rmd
Fixed about 80% of the mess. Still fixing the rest

---
title: "How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach"
Author: "Fiona Ganie"  
Student: "GNXFIO001"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 
    fig_height: 4
abstract: |
  The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has gained popularity in the exchange rate market due to its ease of implementation and tractability. With the evolution of computational power, soft computing techniques have since been used in exchange rate markets. The ARIMA model has been used as the standard benchmark model against which these complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. Prophet is a sophisticated model that differs to traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation. This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis. It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

```

# Introduction \label{Introduction}
The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009]. Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996]. This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability. The ARIMA model is easy to use and yields good forecasts over short forecast horizons, however the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].

With the evolution of computational power, non-linear, soft computing techniques have been proposed as a solution. The ARIMA model has been used as the standard benchmark model against which these more complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. The ARIMA model is tractable and less computationally expensive. It has been used as the building blocks for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.

In December 2016, Facebook open-sourced their forecasting model Prophet. Prophet differs to traditional time series model such as the ARIMA, in that it can produce high quality forecasts in a straightforward way. Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.  It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. This knowledge can be easily incorporated into the model through its intuitively adjustable parameters [@taylor2017]. Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017]. This allows analysts without knowledge on how to pre-process the data, to utilise the model without sacrificing predictive accuracy.  If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.

This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using a Mincer-Zarnowitz approach of measuring forecast accuracy. More specifically, this paper aims to:

1. Compare the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.

2. Compare the out-of-sample forecast performance with missing values and outliers present in the data.

Section \ref{Background and Theory} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model. Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper. Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy. The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}. Finally, Section \ref{Discussion and Conclusions discusses the results in a statistical context.

# Background and Theory \label{Background and Theory}

## Forecasting Exchange Rates with the ARIMA model
ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997]. The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals. The weighted average of the past p values of the process represents an autoregressive (AR) process of order p. It feeds back past values of the process into the current value, inducing correlation between all lags of the process. The weighted average of the q lagged residuals represents a moving average (MA) process of order q. The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process. The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:

@@

The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes. The parameter d is the degree of differencing required to render the process stationary. If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary. The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$. The ARIMA model has the form:

@@

ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012]. They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time. @kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models. In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN) to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model. Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons. Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates. When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike’s Information Criterion (AIC).

The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001]. Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data. Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings. The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing. @etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality. Their results showed that the SARIMA model adequately described the variation in the exchange rate series. The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain. The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993]. This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993]. @cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.

Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009]. Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.

## Other techniques used to forecast exchange rates
Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques. These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002]. Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002]. Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods. They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011]. In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates. Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance. Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009]. Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided. @santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model. He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts. By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced. Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998]. Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple. Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.

## Hybrid ARIMA Models
Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models. By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations. Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.

The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001]. In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain. On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data. The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations. @tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power. The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.

Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006]. In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used. The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties. @ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates. When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.

Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory. They are simple models that are easy to implement and have a consistent method of model design and selection. ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting. The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.

## Forecasting with Prophet 
The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series. Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process. Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared. This is Facebook’s definition of forecasting at scale.
Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017]. The decomposable model is of the form:

@@

Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions. This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly. The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017].

The growth component of Prophet may be modelled as a linear or non-linear function of time. Linear growth is modelled by a piecewise constant function while non-linear growth is modelled similar to population growths which use a logistic growth model [@taylor2017]. A time-varying upper limit may be specified for logistic growth, at which point the forecasts will saturate. This carrying capacity allows the analyst to incorporate their prior knowledge about the maximum obtainable growth level such as the total market or population size into the model. Prophet accounts for changes in the trajectory of this trend by automatically detecting and selecting changepoints in the data at which the growth rate is allowed to change. These changepoints have a Laplace prior distribution placed on them and its scale parameter may be used to adjust the flexibility of the trend and to choose how aggressively the model should follow historical trend changes. Analysts may also adjust the number of potential changepoints included or manually specify their location. This allows non-experts with knowledge about events that may affect the growth rate to use the parameter as a knob to either increase or decrease the number of changepoints included [@taylor2017]. It also allows for the analyst to add changepoints which the automatic selection procedure may have missed or remove changepoints when the model is overfitting historical trends [@taylor2017].

@@

Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model. @taylor2017 forecasted the number of events on Facebook using Prophet. The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend. The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naïve, and the naïve model. While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile. No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.

@@

The data set used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period. The data was obtained from Bitcoincharts.

Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets. The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority (Fantazzini and Nigmatullin 2016).  In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time (Fantazzini and Nigmatullin 2016). Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.

# Methodology \label{Methodology}
In this paper, the out-of-sample forecasts of ARIMA and Prophet are compared over varying forecast horizons. The data is first split into a training set and test set. The training set starts on the 24th January 2016 and ends on the 16th January 2017, while the test set starts on the 17th January 2017 and ends on the 17th July 2017. The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet’s semi-automatic procedure is used for model selection. The selected ARIMA and Prophet model are then fitted to the training set and rolling window forecasts are made 1 day, 30 days and 90 days ahead. This allows us to examine the forecast horizon effect. The Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set. Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Mean Squared Error (MSE) is employed. The results obtained from the ARIMA model will be used as a benchmark for comparison. This process is then repeated based on data which has outliers and missing values present.

## The Box-Jenkins Methodology 
Box and Jenkins proposed a set of guidelines that can be followed when selecting ARIMA models. This systematic procedure to designing ARIMA models has made them highly popular [@spyros1997]. It consists of a four-stage iterative process in which:

1. The process is either transformed or differenced to de-trend and stabilise the variance of the data.

2. The autocorrelation and partial autocorrelation plots are used to determine the order of p and q. 
3. The parameters of the model are then estimated.

4. A diagnostic check is performed to ensure that the residuals are a white noise process.

If the residuals are not white noise, steps 2-4 are repeated until a satisfactory model is identified. On the contrary, if the diagnostic check shows that the residuals are random, the developed model will be the final model used for forecasting.

@@

* The trend is set to be linear
* The width of the uncertainty intervals is set to 80% 
* Weekly and yearly seasonality are automatically detected and included in the model if present 
* The smoothing parameter for holidays and seasonality is set at 10 
* The smoothing parameter for the trend is set to 0.05
* The number of potential changepoints is set to 25

This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model. This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week, due to fluctuations in trading activity. Furthermore, a linear trend is appropriate since Bitcoin does not have an upper limit on its closing prices.

>>>>

---
title: "A comparison between the latest Prophet forecasting model and the ARIMA model: An evaluation of BTC/ZAR predictability"
Author: "Fiona Ganie"  
Student: "GNXFIO001"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 4 
    fig_height: 3
abstract: |
  The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has gained popularity in the exchange rate market due to its ease of implementation and tractability. With the evolution of computational power, soft computing techniques have since been used in exchange rate markets. The ARIMA model has been used as the standard benchmark model against which these complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. Prophet is a sophisticated model that differs from traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation. This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis. It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 4, fig.height = 3, fig.pos="H", fig.pos = 'H')

options(knitr.table.format = "latex")
```

# Introduction \label{Introduction}
The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that have attracted attention in financial market forecasting [@khashei2009]. Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996]. This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability. The ARIMA model is easy to use and yields good forecasts over short forecast horizons, however the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].

With the evolution of computational power, non-linear, soft computing techniques have been proposed as a solution. The ARIMA model has been used as the standard benchmark model against which these more complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. The ARIMA model is tractable and less computationally expensive. It has been used as the building block for more advanced models and has provided the inspiration for hybrid versions of the model which have been used in exchange rate forecasting.

In December 2016, Facebook open-sourced their forecasting model Prophet. Prophet differs from traditional time series model such as the ARIMA in that it can produce high quality forecasts in a straightforward way. Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.  It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. This knowledge can easily be incorporated into the model through its intuitively adjustable parameters [@taylor2017]. Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017]. This allows analysts without knowledge on how to pre-process the data, to utilise the model without sacrificing predictive accuracy.  If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.

This paper broadly aims to compare the forecasts produced by the traditional ARIMA model and Prophet through an evaluation of the exchange rate between Bitcoin and the Rand (BTC/ZAR), using the Model Confidence Set and a Mincer-Zarnowitz approach of measuring forecast accuracy. More specifically, this paper aims at:

1. Comparing the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.

2. Comparing the out-of-sample forecast performance with missing values and outliers present in the data.

Section \ref{Background} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting. Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper. Section \ref{Forecasting Methods} explains and compares the formulation of the ARIMA and Prophet model. Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models, the Model Confidence Set and the Mincer-Zarnowitz approach of measuring forecast accuracy. The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}. Finally, section \ref{Discussion and Conclusions} discusses the results and its financial implications in a statistical context.

# Background \label{Background}

## Forecasting Exchange Rates with the ARIMA model
ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012]. The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001]. They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time and have yielded satisfactory results when predicting exchange rates. @nwankwo2014 forecasted the exchange rate between the Naira and dollar, using Akaike's Information Criterion (AIC) as a measure of performance. Diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data.

Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data. Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings. Some of the most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA (ARFIMA) model.

The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing. @etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality. Their results showed that the SARIMA model adequately described the variation in the exchange rate series.

The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain. The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993]. This makes the model parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993]. @cheung1993 fitted the ARFIMA model to examine five exchange rates and found that there was strong evidence of long memory in the exchange rate time series.

## Other techniques used to forecast exchange rates
Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed due to the failure of ARIMA models to capture the volatility in financial markets  [@anastasakis2009]. @fahimifard2009 applied a GARCH model to the Rial/USD and Rial/EUR to compare the forecasts yielded by the GARCH and ARIMA model over varying forecast horizons.  Their results illustrated that the GARCH model outperformed the ARIMA model when using the Root Mean Square Error (RMSE), Mean Square Error (MSE) and Mean Absolute Deviation (MAD) as a performance criteria.

Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques. These complex techniques are non-linear and can fit complex time series more easily [@castillo2002]. Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002]. Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) mimic the structure of the brain and have had many successful applications in forecasting exchange rates. They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011]. In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates. Superior results were produced when compared to the ARIMA and GARCH model using the RMSE, MSE and MAD as a measure of performance.

Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms and have successfully been used in financial forecasting [@khashei2009]. Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided. @santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model. He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one-step ahead forecasts. By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced. Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model and a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model. The performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998]. Like ANNs, there is no systematic approach for designing FLSs and they are only understandable when simple. Although FLSs has the advantage over ARIMA models in that they can be applied to data with few observations available, it yields acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.

The traditional ARIMA model produces less superior forecasts than its hybrid forms and other complex non-linear techniques, however its forecasts are still satisfactory. It is a simple model that is easy to implement and has a consistent method of model design and selection. ARIMA models are more robust and efficient than complex structural models in relation to short-run forecasting. The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook's forecasting method, Prophet, that was recently released.

## Forecasting with Prophet
The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series. Prophet differs from traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process. Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared. This is Facebook’s definition of forecasting at scale.

Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model. @taylor2017 forecasted the number of events on Facebook using Prophet. The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend. The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naive, and the naive model. While exponential smoothing and the seasonal naive model were quite robust, the ARIMA forecasts were fragile. No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations. If Prophet produces forecasts that are as good as the ARIMA model when forecasting BTC/ZAR, it can be compared to more complex forecasting methods that are less tractable such as the hybrid models and machine learning techniques seen earlier.

The dataset used in this study consists of the daily closing prices of BTC/ZAR over the period 24 January 2016 to 17 July 2017. This comprises of a total of 541 trading days and was the chosen time period for analysis due to constraints in obtaining data over a longer period. The data was obtained from Bitcoincharts.

Bitcoin is of specific interest as it presents an interesting parallel to traditional exchange rate markets. The cryptocurrency is built on a decentralised system and as a result its value cannot be directly influenced by a central authority [@fantazzini2016]. In addition, the Bitcoin market has attracted attention worldwide and is currently the leading cryptocurrency, with awareness and adoption of the currency growing over time [@fantazzini2016]. Its novelty makes it a highly volatile hence speculative market and provides an opportunity for forecasting.

# Forecasting Methods \label{Forecasting Methods}
## The ARIMA model
The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals. The weighted average of the past p values of the process represents an autoregressive (AR) process of order p. It feeds back past values of the process into the current value, inducing correlation between all lags of the process. The weighted average of the q lagged residuals represents a moving average (MA) process of order q. The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process. The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:

@@

The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes. The parameter d is the degree of differencing required to render the process stationary. If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary. The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$. The ARIMA model has the form [@hyndman2014]:

@@

## The Prophet model 
Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017]. The decomposable model is of the form:

@@

Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions. This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly. The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017]. 
The growth component of Prophet may be modelled as a linear or non-linear function of time. Linear growth is modelled by a piecewise constant function while non-linear growth is modelled similar to population growths which use a logistic growth model [@taylor2017]. A time-varying upper limit may be specified for logistic growth, at which point the forecasts will saturate. This carrying capacity allows the analyst to incorporate their prior knowledge about the maximum obtainable growth level such as the total market or population size into the model. Prophet accounts for changes in the trajectory of this trend by automatically detecting and selecting changepoints in the data at which the growth rate is allowed to change. These changepoints have a Laplace prior distribution placed on them and its scale parameter may be used to adjust the flexibility of the trend and to choose how aggressively the model should follow historical trend changes. Analysts may also adjust the number of potential changepoints included or manually specify their location. This allows non-experts with knowledge about events that may affect the growth rate to use the parameter as a knob to either increase or decrease the number of changepoints included [@taylor2017]. It also allows for the analyst to add changepoints which the automatic selection procedure may have missed or remove changepoints when the model is overfitting historical trends [@taylor2017].

@@

@@

# Methodology \label{Methodology}
In this paper, the out-of-sample forecasts of ARIMA and Prophet are compared over varying forecast horizons. The data is first split into a training set and test set. The training set starts on the 24th January 2016 and ends on the 16th January 2017, while the test set starts on the 17th January 2017 and ends on the 17th July 2017. The Box-Jenkins methodology is then used to select the correct ARIMA model while Prophet’s semi-automatic procedure is used for model selection. The selected ARIMA and Prophet model are then fitted to the training set and rolling window forecasts are made 1 day, 30 days and 90 days ahead. This allows us to examine the forecast horizon effect. The Model Confidence Set and Mincer-Zarnowitz test is then used to evaluate the forecasts produced by ARIMA and Prophet against the test set. Since there is no consensus on which accuracy statistic best measures the performance of forecasting techniques, the most popular criteria, the Root Mean Squared Error (RMSE) is employed. The results obtained from the ARIMA model will be used as a benchmark for comparison. This process is then repeated based on data which has outliers and missing values present.

## The Box-Jenkins Methodology 
@box1970 proposed a set of guidelines that can be followed when selecting ARIMA models. It consists of a four-stage iterative process in which:

1. The process is either transformed or differenced to de-trend and stabilise the variance of the data. 
2. The autocorrelation and partial autocorrelation plots are used to determine the order of p and q. 
3. The parameters of the model are then estimated.
4. A diagnostic check is performed to ensure that the residuals are a white noise process.

If the residuals are not white noise, steps 2-4 are repeated until a satisfactory model is identified. On the contrary, if the diagnostic check reveals that the residuals are random, the developed model will be the final model used for forecasting.

@@

* The trend is set to be linear.
* The width of the uncertainty intervals is set to 80%. 
* Weekly and yearly seasonality are automatically detected and included in the model if present 
* The smoothing parameter for holidays and seasonality is set at 10 while the smoothing parameter for the trend is set to 0.05.
* The number of potential changepoints is set to 25.

This paper makes use of Prophet’s default settings to fit the data, however weekly seasonality is included in the model. This allows us to account for our prior knowledge about the closing prices of Bitcoin which tend to be lower around the weekend however higher mid-week due to fluctuations in trading activity. Furthermore, a linear trend is appropriate since Bitcoin does not have an upper limit on its closing prices.

## Rolling Window Forecasts
Rolling window forecasts are useful in evaluating the robustness of a forecasting method. In a rolling window forecast, the forecasts are made h-steps at a time and the actual observation rather than the predicted value is used for the next prediction in the forecast horizon. In this way, a poor forecast will not have negative consequences on future forecasts to be made since the observed values will be used to correct itself for the remaining forecasts [@zhang1998]. In this paper, the ARIMA and Prophet model is first fitted to the training set and the returns are forecasted h-days ahead for h = 1, 30, 90. The size of the training set is then increased by one observation and the models are refitted. The next h-day ahead return is forecasted and this process is repeated until all forecasts have been made into the test set. The h-day ahead forecasts are then compared to the h-day ahead observed values in the test set and the forecast errors are calculated to determine the RMSE.

## Model Confidence Set
The Model Confidence Set (MCS) procedure developed by @hansen2011 is a sequential procedure which starts with an initial set of models $M_0$ of size m. The worse model is then eliminated at each step until the null hypothesis of equal predictive ability (EPA) fails to be rejected for all the models belonging to the set at a given confidence level $1 - \alpha$. This smaller set of models $M^*$ is the superior set of models (SSM) with the best case occurring when the SSM consists of a single best model.
